---
title: "Time_Series_Assignment"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tseries)
library(forecast)
class(gas)
```

class of dataset is "ts", which refer to timeseries

```{r}
frequency(gas)
```

The Periodicity of Dataset is 12(Monthly)

```{r}
gas
```

Timeseries Data is available from January-1956 to August-1995

```{r}
plot(gas)
abline(reg = lm(gas~time(gas)))
```
```{r}
boxplot(gas~cycle(gas),xlab = "Month",ylab = "Australian Gas Production")
```

```{r}
library(forecast)
seasonplot(gas,col = rainbow(12))
```

###Observation:

>The Gas production increase over time with each year which may be indicative of an increasing linear trend.

>In the boxplot there is more production of Gas during months May to August with higher means and higher variances than the other months, indicating seasonality with a apparent cycle of 12 months. 

###Data Stationarity Check

The Australian Gas Production dataset in not stationary, becasue the abline indicate bothe the mean and variance are not constant. To validate this, KPSS test can be Performed

H0: Series is Stationary
Ha: Series is not stationary

```{r}
kpss.test(gas)
```

The P-Value from this test is 0.01 and we reject Null hypothesis, which indicates that the series is not stationary.

To make the series stationry,some transformation functions are used

>Log transformation to variance constant

>Differencing can be used to make mean as constant over time.

```{r}
#plot(diff(log(gas)))
plot(diff(diff(log(gas))))
abline(h=0)
```

```{r}
kpss.test(diff(diff(log(gas))))
```


The P-Value from this test is 0.1 and we accept Null hypothesis, which indicates that the series is stationary.


### ARIMA Model

####PACF Plot

To determine how many previous periods required to forecast next peiods

```{r}
pacf(diff(diff(log(gas))))
```

P=1

### ACF plot

How many previous errors are impacting the value to be forecasted.

```{r}
acf(diff(diff(log(gas))))
```
q=2


```{r}
arima_model=arima(log(gas),c(1,2,2),seasonal = list(order=c(1,2,2),period=12))
summary(arima_model)
```
```{r}
arima_pred<-forecast(arima_model,12)
plot(forecast(arima_pred,12))
```

```{r}
accuracy(arima_pred)
```




